---
title: "PML_Assignment_part1"
author: "Ilya Zemskov"
date: "19OCT2015"
output: html_document
---

# Weight Lifting Exersize Research
The puprose of this research is to analize data about Weight Lifting Exercise from http://groupware.les.inf.puc-rio.br/har and deduce wether subjects did Exercises correct or incorrect.
Data was recieved from the accelerometers on the belt, forearm, arm, and dumbell of 6 participants.

## Initializing environment
```{r}
train <- read.csv("pml-training.csv")
test <- read.csv("pml-testing.csv")
library(caret)
library(ggplot2)
library(Amelia)
```

## Preparing data
Since there are many variables in the dataset (160), let's first try to reduce their quantity with those, whose variance is near-zero.
```{r}
nzv <- nearZeroVar(train[, 1:159])
train_nonzv <- train[, -nzv]
```
The size of the remaining dataset is:
```{r, echo=FALSE}
dim(train_nonzv)
```
Let's have a look at the map of NA's vs other values of the variables.
```{r}
missmap(train_nonzv, main="Map of NAs", col=c("yellow", "black"), legend=FALSE)
```
From the map of NAs we can see that there is still many variables with the most values equal to 'NA'
```{r}
findNAColumns <- function(x) {
  flag = 0
  
  countNA <- sum(is.na(x))
  na_ratio <- countNA/length(x)
  # print(na_ratio, zero.print = ".")
  if (na_ratio > 0.9) {
    flag = 1
  }
  flag
}
# Applying our function to the train_nonzv dataset
isTotalNa <- sapply(train_nonzv[, 1:99], findNAColumns)
# Selectin columns from train_nonzv, which contains NA values less then on 90%
train_nonzv_nona <- train_nonzv[, -which(as.logical(isTotalNa))]
```

Reapeting the mismap function on a new dataset...
```{r}
missmap(train_nonzv_nona, main="Map of NAs", col=c("yellow", "black"), legend=TRUE)
```

Now variables contain only minimum NAs
Now, after some predictor reduction let's look is there any correlated predictors.
```{r}
train_for_cor <- train_nonzv_nona[, 5:58]
train_for_cor <- train_for_cor[, which(sapply(train_for_cor, is.numeric))]
corMatrix <- cor(train_for_cor)
# finding correlated predictors
```

Summary of the correation matrix
```{r}
summary(corMatrix[upper.tri(corMatrix)])
```

```{r}
highlyCorPredictors <- findCorrelation(corMatrix, cutoff = 0.75)
nocor <- train_for_cor[, -highlyCorPredictors]
```

Summary of the new corMatrix
```{r}
corMatrix_new <- cor(nocor)
```

Is there any variables that are linear combinations of the other?
```{r}
combo_info <- findLinearCombos(nocor)
```

After all pre-processing steps finally this is the resulting dataset:
```{r}
result_train <- cbind(nocor[, 2:33], train$classe)
colnames(result_train)[33] <- "classe"
```

## Fitting the model

First I split training set 'result_train' on two smaller training sets to be able to estimate out-of-sample error
```{r}
set.seed(111)
inTrain <- createDataPartition(y=result_train$classe, p=0.7, list=F)
rtrain1 <- result_train[inTrain, ]
rtrain2 <- result_train[-inTrain, ]
```

Let't try the 'Random Forest' metod tot rain out algorithm
```{r}
# use 3-fold CV to select optimal tuning parameters
fitControl <- trainControl(method="cv", number=3, verboseIter=F)

# fitting the model for training set
fit <- train(classe ~ ., data=rtrain1, method="rf", trControl=fitControl)

```

Now I use my model to predict the classe values in the validation dataset rtrain2
```{r}
predict2 <- predict(fit, newdata = rtrain2)

# show confusionMatrix
confusionMatrix(rtrain2$classe, predict2)
```

As we can see on the validating dataset, the accuracy of the model is 99.3%.
This is very good result and therefore I will use the 'Random Forest' Algorithm on the Test set.

## Retraining algorithm on the full training dataset

```{r}
fit_full <- train(classe ~ ., data=result_train, method="rf", trControl=fitControl)
```

## Predict the values of the classe variale for the test dataet

```{r}
# preparing test 

# excluding near zero variance vars
test_nonzv <- test[, -nzv]

# excluding NA vars
isTotalNa_test <- sapply(test_nonzv[, 1:99], findNAColumns)
test_nonzv_nona <- test_nonzv[, -which(as.logical(isTotalNa_test))]

#excluding corelated vars
test_for_cor <- test_nonzv_nona[, 5:58]
test_for_cor <- test_for_cor[, which(sapply(test_for_cor, is.numeric))]
nocor_test <- test_for_cor[, -highlyCorPredictors]

# resulting test set
result_test <- nocor_test[, 2: 33]

# making a prediction
predict_for_test <- predict(fit_full, newdata = result_test)

# converting predictions to character format
predict_for_test_ch <- as.character(predict_for_test)

```

# Writing results

```{r}
# create function to write predictions to files
pml_write_files <- function(x) {
    n <- length(x)
    for(i in 1:n) {
        filename <- paste0("problem_id_", i, ".txt")
        write.table(x[i], file=filename, quote=F, row.names=F, col.names=F)
    }
}

# create prediction files to submit
pml_write_files(predict_for_test_ch)
```